---
title: "SpringBoard Traning - Data Warngling Exercise 2 - Titanic Data Set"
author: "Naga"
date: "25 September 2016"
output: html_document
---

####Load the required libraries
```{r}
library("dplyr")
library("tidyr")


```

####Load the original data set and convert to dply tbl_df data frame

```{r}

titanic <- read.csv("D:/Work/Data Science Projects/SpringBoardTraining/Data Wrangling Exercise 2 Dealing with missing values/4. Analysis/R_Scripts/Data/titanic_original.csv", header = TRUE, sep = ",", quote = "\"",blank.lines.skip = TRUE, strip.white = TRUE,na.strings=c("NA","NaN", ""), stringsAsFactors = FALSE, skipNul = TRUE)

titanic <- tbl_df(titanic)
head(titanic)

```
####removing rows with all columns NA
```{r}
titanic <- titanic[!(rowSums(is.na(titanic)) == length(colnames(titanic))),]
```


####check for NA in Embarked column and replace with S
```{r}
dplyr::filter(titanic, is.na(embarked))
titanic$embarked[is.na(titanic$embarked)] <- 'S'
dplyr::filter(titanic, is.na(embarked))
```

####Check missing values in Age column and replace with mean
Could use Median/Mea to replace missing values (https://www.youtube.com/watch?v=2PrgB_O1T0U)
in this case, mean(29.8813) and median(28) are close enough, hence using mean
```{r}
dplyr::filter(titanic, is.na(age))
titanic$age[is.na(titanic$age)] <- mean(titanic$age, na.rm = TRUE)
dplyr::filter(titanic, is.na(age))
```

####check for NA in boat column and replace with "None"
```{r}
dplyr::filter(titanic, is.na(boat))
titanic$boat[is.na(titanic$boat)] <- 'None'
dplyr::filter(titanic, is.na(boat))
```

####unsure of filling in missing cabin number hence creating binary column for cabin number
```{r}
titanic <- titanic %>% rowwise() %>% mutate(has_cabin_number = ifelse(is.na(cabin),0,1))
```

####write the clean file to titanic_clean.csv
```{r}
write.csv(titanic, "D:/Work/Data Science Projects/SpringBoardTraining/Data Wrangling Exercise 2 Dealing with missing values/4. Analysis/R_Scripts/Data/titanic_clean.csv", quote = TRUE, row.names = FALSE)
```

